---
title: 'COVID-19 LA County City-level Analysis'
author: "David Conti"
date: "`r format(Sys.time(), '%d %B %Y')`"

output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(corrplot)
library(ggplot2)
library(summarytools)
library(stargazer)
library(RColorBrewer)

setwd("~/Google Drive/My Drive/Projects/COVID19/LACountyDataReportsApp/Analysis")

august.date <- "2021-08-18"
january.date <- "2021-01-07"
```

# The Data.
City-level information pulled from the following sources:

LA Almanac (race, age, and income)
http://www.laalmanac.com/population/po38.php
http://www.laalmanac.com/employment/em12.php

LA County vaccine percent by city
http://publichealth.lacounty.gov/media/coronavirus/vaccine/vaccine-dashboard.htm

LA County age adjusted incidence rates
http://dashboard.publichealth.lacounty.gov/covid19_surveillance_dashboard/

Data are incidence rates (proportions), vaccination rates (proportions), and characteristics by city.

# Hypotheses:
All hypothesis are on the city-level and care must be taken when interpreting the results as they describe city-level variation and not individual level variation or co-variation.

**Investigation of the December 2020-January 2021 peak for COVID 19 captured by a single day for incidence rates: `r january.date`:**

1. How do race/ethnicity, age, and income describe variation in incidence rates? 

2. Conditioning on race/ethnicity, age, and income, do August vaccination rates (for `r august.date`) capture any variation in incidence rates during the December 2020-January 2021 peak for COVID 19? 

**Investigation of the August 2021 peak for COVID 19 captured by a single day for incidence rates: `r august.date`:**

3. How do August vaccination proportions, race/ethnicity, age, income, and incidence rates for the Dec.-Jan. peak describe variation in incidence rates during the August peak for COVID 19?


```{r Input and Merge Data, include=F}
d.vax <- read.table("covid_vaccine_percent_by_city.csv", header=T, sep=",")
d.vax.aug <- d.vax[d.vax$Date==august.date,]

d.ir <- read.table("covid_aair_by_city.csv", header=T, sep=",")
d.ir.jan <- d.ir[d.ir$Date==january.date,]
d.ir.aug <- d.ir[d.ir$Date==august.date,]

d.race <- read.table("almanac_race_by_city.csv", header=T, sep=",")
d.age <- read.table("almanac_age_group_by_city.csv", header=T, sep=",")
d.income <- read.table("city_income.csv", header=T, sep=",")
d.income$city <- unlist(lapply(d.income$city, FUN=function(v) { strsplit(v, "†") }))

d.ir.m <- merge(d.ir.jan, d.ir.aug, by="city", all=F)
d <- merge(d.ir.m, d.vax.aug, by="city", all=F)
d <- merge(d, d.race, by="city", all=F)
d <- merge(d, d.age, by="city", all=F)
d <- merge(d, d.income, by="city", all=F)

```

# Variable Distributions and Coding {.tabset}
## Cities
```{r cities table }
city.populations <- data.frame(d$city, d$population.x)
names(city.populations) <- c("City", "Population_Size")
city.populations <- city.populations[order(city.populations$Population_Size, decreasing=T), ]
kable(city.populations, format = "simple", align = 'c', caption = "Cities and Population Size")

City <- d$city

```

## Incidence Rates {.tabset}
### January Peak: 2021-01-07
```{r january peak ir }
Jan.IR <- d$adj_case_14day_rate.x
Jan.IR.cat <- relevel(as.factor(ifelse(Jan.IR > quantile(Jan.IR, probs=.75), "High",
                      ifelse(Jan.IR < quantile(Jan.IR, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(Jan.IR, breaks=seq(from=0, to=round(max(Jan.IR),-3), by=200), col="red", main="January Incidence Rates for Cities", xlab="Incidence Rates")
abline(v=quantile(Jan.IR, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

### August Peak: 2021-08-18
```{r august peak ir }
Aug.IR <- d$adj_case_14day_rate.y

hist(Aug.IR, breaks=seq(from=0, to=round(max(Aug.IR),-2), by=100), col="red", main="August Incidence Rates for Cities", xlab="Incidence Rates")

```

## Percent Vaccinated in Each City
```{r percent vaccinated }
Aug.Vax.percent <- as.numeric(unlist(strsplit(d$dose1_all_c_prcent, "%")))
Aug.Vax.cat <- relevel(as.factor(ifelse(Aug.Vax.percent > quantile(Aug.Vax.percent, probs=.75), "High",
                      ifelse(Aug.Vax.percent < quantile(Aug.Vax.percent, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(Aug.Vax.percent, breaks=seq(from=0, to=100, by=5), col="blue", main="Percent Vaccination Rates for Cities", xlab="Percent Vaccinated")
abline(v=quantile(Aug.Vax.percent, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

## Race/Ethnicity {.tabset }
### Proportion of Hispanic Individuals
```{r hispanic}
CityHispanic <- d$Hispanic
CityHispanic.cat <- relevel(as.factor(ifelse(d$Hispanic > quantile(d$Hispanic, probs=.75), "High",
                      ifelse(d$Hispanic < quantile(d$Hispanic, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(CityHispanic, breaks=seq(from=0, to=100, by=5), main="Proportion of Hispanic Individuals", xlab="Proportion")
abline(v=quantile(CityHispanic, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

### Proportion of Black Individuals
```{r black}
CityBlack <- d$Black
CityBlack.cat <- relevel(as.factor(ifelse(d$Black > 10, "High","Low")), "Low")


hist(CityBlack, breaks=seq(from=0, to=100, by=5), main="Proportion of Black Individuals", xlab="Proportion")
abline(v=10, lwd=2, lty=2)
text(10, 60, "Cutoff at 10%", pos=4)
```

### Proportion of Asian Individuals
```{r asian}
CityAsian <- d$Asian
CityAsian.cat <- relevel(as.factor(ifelse(d$Asian > quantile(d$Asian, probs=.75), "High",
                      ifelse(d$Asian < quantile(d$Asian, probs=.5), "Low",
                             "Med"))), ref="Low")

hist(CityAsian, breaks=seq(from=0, to=100, by=5), main="Proportion of Asian Individuals", xlab="Proportion")
abline(v=quantile(CityAsian, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

## Age
Age data is captured by the number/proportion of individuals within each city in a age category. Age categories include: "Age.Under.15","Age.15.17","Age.18.24","Age.25.34","Age.35.54","Age.55.64","Age.65."

Three groups are created for the analysis:

- Young: combined age categories for "Age.Under.15","Age.15.17","Age.18.24"

- Middle: age category for "Age.35.54"

- Old: combined age categories for "Age.55.64","Age.65."

```{r age}
Age <- d[,c("Age.Under.15","Age.15.17","Age.18.24","Age.25.34","Age.35.54","Age.55.64","Age.65.")]
Age <- t(apply(Age, 1, FUN=function(v) { v/sum(v) }))

CityAgeYoung <- apply(Age[,c("Age.Under.15","Age.15.17","Age.18.24")], 1, sum)
CityAgeOld <- apply(Age[,c("Age.55.64","Age.65.")], 1, sum)
CityAgeMid <- Age[,"Age.35.54"]
CityAge <- cbind(CityAgeYoung, CityAgeOld)

CityAgeOld.cat <- relevel(as.factor(ifelse(CityAgeOld > quantile(CityAgeOld, probs=.75), "High","notHigh")), "notHigh")
CityAgeYoung.cat <- relevel(as.factor(ifelse(CityAgeYoung > quantile(CityAgeYoung, probs=.75), "High","notHigh")), "notHigh")

ftable(CityAgeOld.cat, CityAgeYoung.cat)
```

## Household Income
```{r household income}
HouseholdIncome <- d$Households
HouseholdIncome.cat <- relevel(as.factor(ifelse(d$Households > quantile(d$Households, probs=.75), "High",
                      ifelse(d$Households < quantile(d$Households, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(HouseholdIncome, breaks=seq(from=0, to=round(max(HouseholdIncome),-2), by=10000), main="Household Income Within Each City", xlab="Income", col="green")
abline(v=quantile(HouseholdIncome, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

# Variables for Analysis {.tabset}

## Distributions
```{r distributions }
CityAge.m <- cbind(CityAgeOld.cat, CityAgeYoung.cat)
CityRaceEthnicty.m <- cbind(CityHispanic.cat, CityBlack.cat, CityAsian.cat)
d.analysis <- data.frame(Aug.IR, Jan.IR.cat, Aug.Vax.cat, CityRaceEthnicty.m, CityAge.m, HouseholdIncome.cat)

summarytools::view(dfSummary(d.analysis, style = 'grid',
                               max.distinct.values = 10, plain.ascii =   FALSE, valid.col = FALSE, headings = FALSE), method = "render")

```


## Correlation
```{r create variables for analysis}
d.num <- data.frame(Aug.IR, as.numeric(Jan.IR.cat), as.numeric(Aug.Vax.cat),
                    as.numeric(CityHispanic.cat), as.numeric(CityBlack.cat), as.numeric(CityAsian.cat), 
                    as.numeric(CityAgeOld.cat), as.numeric(CityAgeYoung.cat),
                    as.numeric(HouseholdIncome.cat))
cormat <- cor(d.num, use="complete.obs", method="pearson")
corrplot(cormat, type="upper", order="original",
          col=brewer.pal(n=8, name="RdYlBu"),
          title = "",
          addCoef.col = "black",
          tl.cex=.5, number.cex=.5)

```


# Pairwise Trend Plots {.tabset}
## August IR vs. January IR
```{r aug.ir vs. jan.ir }
coef <- summary(lm(Aug.IR~Jan.IR))$coef
ggplot(d.analysis, aes(x=Jan.IR, y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=3000, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=3000, y=800, label=paste("p-value", "=", round(coef[2,4],6)))+
  xlab("January Incidence Rates")+ylab("August Incidence Rates")
```

## August IR vs. August Percent Vaccinated
```{r aug.ir vs. aug vaccination percent } 
coef <- summary(lm(Aug.IR~Aug.Vax.percent))$coef
ggplot(d.analysis, aes(x=Aug.Vax.percent, y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=80, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=80, y=800, label=paste("p-value", "=", round(coef[2,4],2)))+
  xlab("August Vaccination Percent")+ylab("August Incidence Rates")
```

## January IR vs. Household Income
```{r jan.ir vs. household income } 
coef <- summary(lm(Jan.IR~as.numeric(relevel(HouseholdIncome.cat, "Low"))))$coef
ggplot(d.analysis, aes(x=as.numeric(relevel(HouseholdIncome.cat, "Low")), y=Jan.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=3000, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=2500, label=paste("p-value", "=", round(coef[2,4],23))) +
  scale_x_continuous(name ="Household Income", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("January Incidence Rates")
```

## August IR vs. Household Income
```{r aug.ir vs. household income } 
coef <- summary(lm(Aug.IR~as.numeric(relevel(HouseholdIncome.cat, "Low"))))$coef
ggplot(d.analysis, aes(x=as.numeric(relevel(HouseholdIncome.cat, "Low")), y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=3000, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=2500, label=paste("p-value", "=", round(coef[2,4],23))) +
  scale_x_continuous(name ="Household Income", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("August Incidence Rates")
```

# January Incident Rates: {.tabset}
## Regression Model {.tabset}
```{r jan.ir regression }
reg.Jan.noVax <- lm(Jan.IR ~ HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)

```

### Regression summary
```{r jan.ir regression: summary }
kable(summary(reg.Jan.noVax)$coef, digits=3, format = "simple", align = 'c', caption = "January Regression")
#stargazer(reg.Jan.noVax, type = 'html', ci=TRUE, ci.level=0.95, single.row = T)
```

### Regression ANOVA
```{r jan.ir regression: anova }
kable(anova(reg.Jan.noVax), digits=3, format = "simple", align = 'c', caption = "January Regression ANOVA")
```

## Regression Model with August Vaccination {.tabset}
```{r jan.ir regression with vaccination}
reg.Jan <- lm(Jan.IR ~ Aug.Vax.cat+HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)

```

### Regression summary
```{r jan.ir regression with vax: summary }
kable(summary(reg.Jan)$coef, digits=3, format = "simple", align = 'c', caption = "January Regression with Vaccination")
```

### Regression ANOVA
```{r jan.ir regression with vax: anova }
kable(anova(reg.Jan), digits=3, format = "simple", align = 'c', caption = "January Regression with Vaccination ANOVA")
```

# August Incident Rates: 
## Regression Model {.tabset}
```{r aug.ir regression }
reg.Aug <- lm(Aug.IR ~ Aug.Vax.cat + Jan.IR.cat + HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)
```

### Regression summary
```{r aug.ir regression: summary }
kable(summary(reg.Aug)$coef, digits=3, format = "simple", align = 'c', caption = "August Regression")
```

### Regression ANOVA
```{r aug.ir regression: anova }
kable(anova(reg.Aug), digits=3, format = "simple", align = 'c', caption = "August Regression ANOVA")
```
