---
title: 'COVID-19 LA County City-level Analysis'
author: "David Conti"
date: "`r format(Sys.time(), '%d %B %Y')`"

output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(corrplot)
library(ggplot2)
library(summarytools)
library(stargazer)
library(RColorBrewer)
library(nnet)
library(lubridate)
library(cowplot)

# setwd("~/Google Drive/My Drive/Projects/COVID19/LACountyDataReportsApp/Analysis")
august.date <- "2021-08-18"
january.date <- "2021-01-07"
```

# The Data.
City-level information pulled from the following sources:

LA Almanac (race, age, and income)
http://www.laalmanac.com/population/po38.php
http://www.laalmanac.com/employment/em12.php

LA County vaccine percent by city
http://publichealth.lacounty.gov/media/coronavirus/vaccine/vaccine-dashboard.htm

LA County age adjusted incidence rates
http://dashboard.publichealth.lacounty.gov/covid19_surveillance_dashboard/

Data are incidence rates (proportions), vaccination rates (proportions), and characteristics by city.

# Hypotheses:
All hypothesis are on the city-level and care must be taken when interpreting the results as they describe city-level variation and not individual level variation or co-variation.

**Investigation of the December 2020-January 2021 peak for COVID 19 captured by a single day for incidence rates: `r january.date`:**

1. How do race/ethnicity, age, and income describe variation in incidence rates? 

**Investigation of the August 2021 peak for COVID 19 captured by a single day for incidence rates: `r august.date`:**

2. How do August vaccination proportions, race/ethnicity, age, income, and incidence rates for the Dec.-Jan. peak describe variation in incidence rates during the August peak for COVID 19?

**Investigation of August 2021 vaccination rates for COVID 19 captured by a single day for vaccination rates: `r august.date`:**

3. How do race/ethnicity, age, income, and incidence rates for the Dec.-Jan. peak describe variation in August vaccination rates?

```{r Input and Merge Data, include=F}

tmp_data <- readRDS("data/analysis_data.rds")

# d.vax <- read.table("data/covid_vaccine_percent_by_city.csv", header=T, sep=",")
d.vax <- read.table("data/LAC_Vaccine_City_Data_Clean.csv", header=T, sep=",")
d.vax.aug <- tmp_data[tmp_data$Date.x==august.date, c('city', 'dose1_all_c_prcent')]

# d.ir <- read.table("data/covid_aair_by_city.csv", header=T, sep=",")
d.ir <- tmp_data
d.ir.jan <- d.ir[d.ir$Date.x==january.date,]
d.ir.aug <- d.ir[d.ir$Date.x==august.date,]

d.race <- read.table("data/almanac_race_by_city.csv", header=T, sep=",")
d.age <- read.table("data/almanac_age_group_by_city.csv", header=T, sep=",")
d.income <- read.table("data/city_income.csv", header=T, sep=",")
d.income$city <- unlist(lapply(d.income$city, FUN=function(v) { strsplit(v, "†") }))

# d.ir.m <- merge(d.ir.aug, d.ir.jan[, c('city', ')], by="city", all=F)
d <- merge(d.ir, d.vax.aug, by="city", all=F)
d <- merge(d, d.race, by="city", all=F)
d <- merge(d, d.age, by="city", all=F)
d <- merge(d, d.income, by="city", all=F)

```

# Variable Distributions and Coding {.tabset}
## Cities
```{r cities table }
d <- readRDS("data/analysis_data.rds") %>% 
  arrange(Date.x) %>% 
  filter(Date.x == "2021-08-18")

city.populations <- data.frame(d$city, d$population.x)
names(city.populations) <- c("City", "Population_Size")
city.populations <- city.populations[order(city.populations$Population_Size, decreasing=T), ]
kable(city.populations, format = "simple", align = 'c', caption = "Cities and Population Size")

City <- d$city

```

## Incidence Rates {.tabset}
### January Peak: 2021-01-07
```{r january peak ir }
Jan.IR <- d$adj_case_14day_rate.y
Jan.IR.cat <- relevel(as.factor(ifelse(Jan.IR > quantile(Jan.IR, probs=.75), "High",
                      ifelse(Jan.IR < quantile(Jan.IR, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(Jan.IR, breaks=seq(from=0, to=round(max(Jan.IR),-3), by=200), col="red", main="January Incidence Rates for Cities", xlab="Incidence Rates")
abline(v=quantile(Jan.IR, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

### August Peak: 2021-08-18
```{r august peak ir }
Aug.IR <- d$adj_case_14day_rate.x

hist(Aug.IR, breaks=seq(from=0, to=round(max(Aug.IR),-2), by=100), col="red", main="August Incidence Rates for Cities", xlab="Incidence Rates")

```

## Percent Vaccinated in Each City
```{r percent vaccinated }
Aug.Vax.percent <- d$dose1_all_c_prcent
Aug.Vax.cat <- relevel(as.factor(ifelse(Aug.Vax.percent > quantile(Aug.Vax.percent, probs=.75), "High",
                      ifelse(Aug.Vax.percent < quantile(Aug.Vax.percent, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(Aug.Vax.percent, breaks=seq(from=0, to=100, by=5), col="blue", main="Percent Vaccination Rates for Cities", xlab="Percent Vaccinated")
abline(v=quantile(Aug.Vax.percent, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

## Race/Ethnicity {.tabset }
### Proportion of Hispanic Individuals
```{r hispanic}
CityHispanic <- d$Hispanic
CityHispanic.cat <- relevel(as.factor(ifelse(d$Hispanic > quantile(d$Hispanic, probs=.75), "High",
                      ifelse(d$Hispanic < quantile(d$Hispanic, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(CityHispanic, breaks=seq(from=0, to=100, by=5), main="Proportion of Hispanic Individuals", xlab="Proportion")
abline(v=quantile(CityHispanic, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

### Proportion of Black Individuals
```{r black}
CityBlack <- d$Black
CityBlack.cat <- relevel(as.factor(ifelse(d$Black > 10, "High","Low")), "Low")


hist(CityBlack, breaks=seq(from=0, to=100, by=5), main="Proportion of Black Individuals", xlab="Proportion")
abline(v=10, lwd=2, lty=2)
text(10, 60, "Cutoff at 10%", pos=4)
```

### Proportion of Asian Individuals
```{r asian}
CityAsian <- d$Asian
CityAsian.cat <- relevel(as.factor(ifelse(d$Asian > quantile(d$Asian, probs=.75), "High",
                      ifelse(d$Asian < quantile(d$Asian, probs=.5), "Low",
                             "Med"))), ref="Low")

hist(CityAsian, breaks=seq(from=0, to=100, by=5), main="Proportion of Asian Individuals", xlab="Proportion")
abline(v=quantile(CityAsian, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

## Age
Age data is captured by the number/proportion of individuals within each city in a age category. Age categories include: "Age.Under.15","Age.15.17","Age.18.24","Age.25.34","Age.35.54","Age.55.64","Age.65."

Three groups are created for the analysis:

- Young: combined age categories for "Age.Under.15","Age.15.17","Age.18.24"

- Middle: age category for "Age.35.54"

- Old: combined age categories for "Age.55.64","Age.65."

```{r age}
Age <- d[,c("Age.Under.15","Age.15.17","Age.18.24","Age.25.34","Age.35.54","Age.55.64","Age.65.")]
Age <- t(apply(Age, 1, FUN=function(v) { v/sum(v) }))

CityAgeYoung <- apply(Age[,c("Age.Under.15","Age.15.17","Age.18.24")], 1, sum)
CityAgeOld <- apply(Age[,c("Age.55.64","Age.65.")], 1, sum)
CityAgeMid <- Age[,"Age.35.54"]
CityAge <- cbind(CityAgeYoung, CityAgeOld)

CityAgeOld.cat <- relevel(as.factor(ifelse(CityAgeOld > quantile(CityAgeOld, probs=.75), "High","notHigh")), "notHigh")
CityAgeYoung.cat <- relevel(as.factor(ifelse(CityAgeYoung > quantile(CityAgeYoung, probs=.75), "High","notHigh")), "notHigh")

ftable(CityAgeOld.cat, CityAgeYoung.cat)
```

## Household Income
```{r household income}
HouseholdIncome <- d$Households
HouseholdIncome.cat <- relevel(as.factor(ifelse(HouseholdIncome > quantile(HouseholdIncome, probs=.75), "High",
                      ifelse(HouseholdIncome < quantile(HouseholdIncome, probs=.25), "Low",
                             "Med"))), ref="Med")

hist(HouseholdIncome, breaks=seq(from=0, to=round(max(HouseholdIncome),-2), by=10000), main="Household Income Within Each City", xlab="Income", col="green")
abline(v=quantile(HouseholdIncome, probs=c(.25, .5, .75)), lwd=2, lty=2)

```

# Variables for Analysis {.tabset}

## Distributions
```{r distributions }
CityAge.m <- cbind(CityAgeOld.cat, CityAgeYoung.cat)
CityRaceEthnicty.m <- cbind(CityHispanic.cat, CityBlack.cat, CityAsian.cat)
d.analysis <- data.frame(Aug.IR, Jan.IR.cat, Aug.Vax.cat, Aug.Vax.percent, CityRaceEthnicty.m, CityAge.m, HouseholdIncome.cat)

summarytools::view(dfSummary(d.analysis, style = 'grid',
                               max.distinct.values = 10, plain.ascii = FALSE, valid.col = FALSE, headings = FALSE), method = "render")

```


## Correlation
```{r create variables for analysis}

# need to make sure low is lowest when setting as.numeric (to make cor more clear)





d.num <- data.frame(Aug.IR, Jan.IR.cat, Aug.Vax.cat,
                    CityHispanic.cat, as.numeric(CityBlack.cat), as.numeric(CityAsian.cat), 
                    as.numeric(CityAgeOld.cat), as.numeric(CityAgeYoung.cat),
                    HouseholdIncome.cat) %>% 
  mutate(across(c(Jan.IR.cat, Aug.Vax.cat, CityHispanic.cat, HouseholdIncome.cat), 
                ~ as.numeric(relevel(.x, ref = "Low"))))

cormat <- cor(d.num, use="complete.obs", method="pearson")
corrplot(cormat, type="upper", order="original",
          col=brewer.pal(n=8, name="RdYlBu"),
          title = "",
          addCoef.col = "black",
          tl.cex=.5, number.cex=.5)

```


# Pairwise Trend Plots {.tabset}
## January IR vs. Proportion of Hispanic Individuals
```{r jan.ir vs. proportion of hispanic individuals } 
CityHispanic.cont <- as.numeric(relevel(CityHispanic.cat, "Low"))
coef <- summary(lm(Jan.IR~CityHispanic.cont))$coef
ggplot(d.analysis, aes(x=CityHispanic.cont, y=Jan.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=3000, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=2500, label=paste("p-value", "=", round(coef[2,4],25))) +
  scale_x_continuous(name ="Proportion of Hispanic Individuals", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("Jan Incidence Rates")
```

## January IR vs. Household Income
```{r jan.ir vs. household income } 
coef <- summary(lm(Jan.IR~as.numeric(relevel(HouseholdIncome.cat, "Low"))))$coef
ggplot(d.analysis, aes(x=as.numeric(relevel(HouseholdIncome.cat, "Low")), y=Jan.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=3000, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=2500, label=paste("p-value", "=", round(coef[2,4],23))) +
  scale_x_continuous(name ="Household Income", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("January Incidence Rates")
```

## August IR vs. January IR
```{r aug.ir vs. jan.ir }
coef <- summary(lm(Aug.IR~Jan.IR))$coef
ggplot(d.analysis, aes(x=Jan.IR, y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=3000, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=3000, y=800, label=paste("p-value", "=", round(coef[2,4],6)))+
  xlab("January Incidence Rates")+ylab("August Incidence Rates")
```

## August IR vs. Proportion of Hispanic Individuals
```{r aug.ir vs. proportion of hispanic individuals } 
CityHispanic.cont <- as.numeric(relevel(CityHispanic.cat, "Low"))
coef <- summary(lm(Aug.IR~CityHispanic.cont))$coef
ggplot(d.analysis, aes(x=CityHispanic.cont, y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=800, label=paste("p-value", "=", round(coef[2,4],2))) +
  scale_x_continuous(name ="Proportion of Hispanic Individuals", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("August Incidence Rates")
```


## August IR vs. Household Income
```{r aug.ir vs. household income } 
coef <- summary(lm(Aug.IR~as.numeric(relevel(HouseholdIncome.cat, "Low"))))$coef
ggplot(d.analysis, aes(x=as.numeric(relevel(HouseholdIncome.cat, "Low")), y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=2.5, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=2.5, y=800, label=paste("p-value", "=", round(coef[2,4],2))) +
  scale_x_continuous(name ="Household Income", breaks=c(1,2,3), labels=c("Low", "Med", "High"))+
  ylab("August Incidence Rates")
```

## August IR vs. August Percent Vaccinated
```{r aug.ir vs. aug vaccination percent } 
coef <- summary(lm(Aug.IR~Aug.Vax.percent))$coef
ggplot(d.analysis, aes(x=Aug.Vax.percent, y=Aug.IR)) +
  geom_point(color="blue", size=2) +
  geom_smooth(method=lm, color="black") +
  geom_text(x=80, y=900, label=paste(expression(beta), "=", round(coef[2,1],2))) +
  geom_text(x=80, y=800, label=paste("p-value", "=", round(coef[2,4],2)))+
  xlab("August Vaccination Percent")+ylab("August Incidence Rates")
```


# January Incident Rates: {.tabset}
## Regression Model {.tabset}
```{r jan.ir regression }
reg.Jan.noVax <- lm(Jan.IR ~ HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)

```

### Regression summary
```{r jan.ir regression: summary }
kable(summary(reg.Jan.noVax)$coef, digits=3, format = "simple", align = 'c', caption = "January Regression")
#stargazer(reg.Jan.noVax, type = 'html', ci=TRUE, ci.level=0.95, single.row = T)
```

### Regression ANOVA
$R^2 =$ `r round(summary(reg.Jan.noVax)$r.squared,2)`
```{r jan.ir regression: anova }
kable(anova(reg.Jan.noVax), digits=3, format = "simple", align = 'c', caption = "January Regression ANOVA")
```

# August Incident Rates: 
## Vaccination Only {.tabset}
### Regression Model {.tabset}
```{r aug.ir vax only regression }
reg.Aug.vax <- lm(Aug.IR ~ Aug.Vax.cat)
```

#### Regression summary
```{r aug.ir vax only regression: summary }
kable(summary(reg.Aug.vax)$coef, digits=3, format = "simple", align = 'c', caption = "August Regression")
```

#### Regression ANOVA
$R^2 =$ `r round(summary(reg.Aug.vax)$r.squared,2)`
```{r aug.ir vax only regression: anova }

kable(anova(reg.Aug.vax), digits=3, format = "simple", align = 'c', caption = "August Regression ANOVA")
```


## Full Model
### Regression Model {.tabset}
```{r aug.ir regression }
reg.Aug <- lm(Aug.IR ~ Aug.Vax.cat + Jan.IR.cat + HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)
```

#### Regression summary
```{r aug.ir regression: summary }
kable(summary(reg.Aug)$coef, digits=3, format = "simple", align = 'c', caption = "August Regression")
```

#### Regression ANOVA
$R^2 =$ `r round(summary(reg.Aug)$r.squared,2)`
```{r aug.ir regression: anova }
kable(anova(reg.Aug), digits=3, format = "simple", align = 'c', caption = "August Regression ANOVA")
```

# August Vaccination Rates:
## Regression Model with August Vaccination {.tabset}
```{r aug.vax.percent regression}
reg.Aug.vax <- lm(Aug.Vax.percent ~ Jan.IR.cat+HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)

```

### Regression summary
```{r aug.vax.percent regression: summary }
kable(summary(reg.Aug.vax)$coef, digits=3, format = "simple", align = 'c', caption = "August Vaccination Regression")
```

### Regression ANOVA
```{r aug.vax.percent regression: anova }
kable(anova(reg.Aug.vax), digits=3, format = "simple", align = 'c', caption = "August Vaccination Regression ANOVA")
```

## Multnomial Regression Model with August Vaccination {.tabset}
```{r aug.vax.percent multinomial regression}
reg.Aug.vax <- lm(Aug.Vax.percent ~ Jan.IR.cat+HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)
reg.Aug.vax <- multinom(Aug.Vax.cat ~ Jan.IR.cat+HouseholdIncome.cat+CityRaceEthnicty.m+CityAge.m)


```

### Regression summary for High Vaccination Rates
```{r aug.vax.percent multinomial regression: high vax summary }
coef.high <- summary(reg.Aug.vax)$coefficients["High",]
OR.high <- exp(coef.high)
se.high <- summary(reg.Aug.vax)$standard.errors["High",]
z.high <- coef.high/se.high
p.value.high <- (1 - pnorm(abs(z.high), 0, 1)) * 2
r.high <- data.frame(OR.high, coef.high, se.high, z.high, p.value.high)
kable(r.high, digits=3, format = "simple", align = 'c', caption = "August Vaccination Regression for High Vax Rates")
```

### Regression summary for Low Vaccination Rates
```{r aug.vax.percent multinomial regression: low vax summary }
coef.low <- summary(reg.Aug.vax)$coefficients["Low",]
OR.low <- exp(coef.low)
se.low <- summary(reg.Aug.vax)$standard.errors["Low",]
z.low <- coef.low/se.low
p.value.low <- (1 - pnorm(abs(z.low), 0, 1)) * 2
r.low <- data.frame(OR.low, coef.low, se.low, z.low, p.value.low)
kable(r.low, digits=3, format = "simple", align = 'c', caption = "August Vaccination Regression for Low Vax Rates")
```


<br>

---

<br>

## Model params over time {.tabset}

Full model fit by date
\
``{r} adj_case_14day_rate.x ~ Vax.percent + Jan.IR.cat + HouseholdIncome.cat+CityHispanic.cat+CityBlack.cat+ CityAsian.cat+ CityAgeOld.cat+ CityAgeYoung.cat`` 
\
No vaccination lag time in this example

### Vaccination rates
Modeled as a continuous variable. Plots include dates > 04/01/2021
```{r}

d <- readRDS("data/analysis_data.rds") 
lac_summarytable_dash <- read.csv("data/LA_County_Covid19_CSA_14day_case_death_table_Clean.csv")

# data clean ----
dout <- d %>% 
  mutate(Date.x = ymd(Date.x), 
         Vax.percent = dose1_all_c_prcent, 
         Jan.IR =  as.numeric(adj_case_14day_rate.y),
         Jan.IR.cat =  relevel(as.factor(ifelse(Jan.IR > quantile(Jan.IR, probs=.75), "High",
                                                ifelse(Jan.IR < quantile(Jan.IR, probs=.25), "Low",
                                                       "Med"))), ref="Low")) %>%
  dplyr::select(Date.x, city, adj_case_14day_rate.x, 
                Vax.percent, Jan.IR.cat, 
                HouseholdIncome.cat, CityHispanic.cat, CityBlack.cat, 
                CityAsian.cat, CityAgeOld.cat, CityAgeYoung.cat)




test <- dout %>% 
  # dplyr::filter(Date.x > "2021-03-01") %>% 
  tidyr::nest(data = -Date.x) %>% 
  dplyr::mutate(fit = purrr::map(data, ~lm(adj_case_14day_rate.x ~ Vax.percent + Jan.IR.cat + HouseholdIncome.cat+CityHispanic.cat+CityBlack.cat+ CityAsian.cat+ CityAgeOld.cat+ CityAgeYoung.cat, data = .x)),
                tidied = purrr::map(fit, ~broom::tidy(.x), quality = purrr::map(fit, ~glance(.x)))) %>% 
  dplyr::select(-data, -fit) %>% 
  tidyr::unnest(tidied) %>% 
  # dplyr::filter(grepl("Vax", term)) %>%
  arrange(Date.x)



# Vaccination rates ---- 
# after april 2021 (vaccination rates too low prior)
p1 <- ggplot(test %>% filter(grepl("Vax.percent", term), Date.x > "2021-04-01"), aes(Date.x, estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))
  # geom_line(data = lac_summarytable_dash %>% filter(city == "Los Angeles County"), aes(Date, adj_case_14day_rate))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County", 
                      Date > "2021-04-01") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')


```

### Race
```{r, warning=F}

p1 <- ggplot(test %>% filter(term %in% c("CityAsian.catHigh", "CityBlack.catHigh", "CityHispanic.catHigh")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```


### Household income
```{r}
p1 <- ggplot(test %>% filter(term %in% c("HouseholdIncome.catHigh", "HouseholdIncome.catLow")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.2))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```



### Age category
```{r}
p1 <- ggplot(test %>% filter(term %in% c("CityAgeOld.catHigh", "CityAgeYoung.catHigh")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```


### January incidence rates
```{r}

p1 <- ggplot(test %>% filter(term %in% c("Jan.IR.catHigh", "Jan.IR.catMed")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))
# geom_line(data = lac_summarytable_dash %>% filter(city == "Los Angeles County"), aes(Date, adj_case_14day_rate))
# ggsave(p1, width = 10, height = 8, file = "~/Desktop/test1.png")
p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```




<br>

---

<br>

## Model params over time (vax lag) {.tabset}

Full model fit by date
\
``{r} adj_case_14day_rate.x ~ Vax.percent + Jan.IR.cat + HouseholdIncome.cat+CityHispanic.cat+CityBlack.cat+ CityAsian.cat+ CityAgeOld.cat+ CityAgeYoung.cat`` 
\
Introduce 14 day lag between vaccination rates and COVID incidence rates

### Vaccination rates
Modeled as a continuous variable. Plots include dates > 04/01/2021
```{r}

d <- readRDS("data/analysis_data_14vax_lag.rds") 
lac_summarytable_dash <- read.csv("data/LA_County_Covid19_CSA_14day_case_death_table_Clean.csv")

# data clean ----
dout <- d %>% 
  mutate(Date.x = ymd(Date.x), 
         Vax.percent = dose1_all_c_prcent, 
         Jan.IR =  as.numeric(adj_case_14day_rate.y),
         Jan.IR.cat =  relevel(as.factor(ifelse(Jan.IR > quantile(Jan.IR, probs=.75), "High",
                                                ifelse(Jan.IR < quantile(Jan.IR, probs=.25), "Low",
                                                       "Med"))), ref="Low")) %>%
  dplyr::select(Date.x, city, adj_case_14day_rate.x, 
                Vax.percent, Jan.IR.cat, 
                HouseholdIncome.cat, CityHispanic.cat, CityBlack.cat, 
                CityAsian.cat, CityAgeOld.cat, CityAgeYoung.cat)




test <- dout %>% 
  # dplyr::filter(Date.x > "2021-03-01") %>% 
  tidyr::nest(data = -Date.x) %>% 
  dplyr::mutate(fit = purrr::map(data, ~lm(adj_case_14day_rate.x ~ Vax.percent + Jan.IR.cat + HouseholdIncome.cat+CityHispanic.cat+CityBlack.cat+ CityAsian.cat+ CityAgeOld.cat+ CityAgeYoung.cat, data = .x)),
                tidied = purrr::map(fit, ~broom::tidy(.x), quality = purrr::map(fit, ~glance(.x)))) %>% 
  dplyr::select(-data, -fit) %>% 
  tidyr::unnest(tidied) %>% 
  # dplyr::filter(grepl("Vax", term)) %>%
  arrange(Date.x)



# Vaccination rates ---- 
# after april 2021 (vaccination rates too low prior)
p1 <- ggplot(test %>% filter(grepl("Vax.percent", term), Date.x > "2021-04-01"), aes(Date.x, estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))
  # geom_line(data = lac_summarytable_dash %>% filter(city == "Los Angeles County"), aes(Date, adj_case_14day_rate))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County", 
                      Date > "2021-04-01") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')


```

### Race
```{r, warning=F}

p1 <- ggplot(test %>% filter(term %in% c("CityAsian.catHigh", "CityBlack.catHigh", "CityHispanic.catHigh")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```


### Household income
```{r}
p1 <- ggplot(test %>% filter(term %in% c("HouseholdIncome.catHigh", "HouseholdIncome.catLow")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.2))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```



### Age category
```{r}
p1 <- ggplot(test %>% filter(term %in% c("CityAgeOld.catHigh", "CityAgeYoung.catHigh")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))

p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```


### January incidence rates
```{r}

p1 <- ggplot(test %>% filter(term %in% c("Jan.IR.catHigh", "Jan.IR.catMed")), aes(Date.x, estimate)) + 
  geom_point(aes(colour = term)) + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error, colour = term)) +
  geom_hline(yintercept = 0) + theme_bw() + theme(legend.position = c(0.8, 0.8))
# geom_line(data = lac_summarytable_dash %>% filter(city == "Los Angeles County"), aes(Date, adj_case_14day_rate))
# ggsave(p1, width = 10, height = 8, file = "~/Desktop/test1.png")
p2 <- ggplot(lac_summarytable_dash %>% 
               filter(city == "Los Angeles County") %>% 
               mutate(Date = ymd(Date)), aes(Date, adj_case_14day_rate)) +
  geom_line()

plot_grid(p1, p2, ncol = 1, align = 'v')

```




